---
title: "FTICR Visualizations"
output: html_notebook
---


```{r}
# Clean workspace
rm(list=ls())

library("ftmsRanalysis")
library("vegan")
library("ggplot2")
library("ape")
library("ggVennDiagram")
library("reshape2")
library("plyr")
```


# Load Peak Object

```{r}
load("ficusPeakObj.Robj")
str(peakObj)
```

```{r}
peakObj
```


### Binary Matrix

```{r}
# Set rownames
edata <- peakObj$e_data
rownames(edata) <- as.character(edata$Mass)
edata$Mass <- NULL
# Binary matrix
edata.binary <- apply(edata,2,function(y)as.numeric(y>0))
# Na's as zero
edata.binary[is.na(edata.binary)] <- 0
# As dataframe edata.binary
edata.binary <- as.data.frame(edata.binary)
# Add Mass to binary matrix
edata.binary$Mass <- rownames(edata)
# Add metadata to binary matrix
tmp <- peakObj$e_meta[,c("Mass","bs1_class")] # save to tmp; direct merge isn't working
str(tmp)
# Merge 
edata.binary <- merge(edata.binary,tmp,by="Mass",all = T)
# Remove Mass column from df
edata.binary$Mass <- NULL
```


```{r}
# Carbon compounds #
carbon.cpds <- aggregate(.~bs1_class,data=edata.binary,FUN=sum)
# Transfor count to relative abundance #
carbon.cpds[,-1] <- decostand(carbon.cpds[,-1],method = "total",MARGIN = 2)*100
# Sum check
colSums(carbon.cpds[,-1])
```

```{r}
# melt df #
carbon.cpds <- melt(carbon.cpds,variable.name = "Sample_ID")
# Add sample info
carbon.cpds <- merge(carbon.cpds,peakObj$f_data,by="Sample_ID")
head(carbon.cpds)
```

```{r,fig.width=15}
# Color palette #
pal12 <- c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C", "#FB9A99","#E31A1C", "#FDBF6F", "#FF7F00", "#CAB2D6", "#6A3D9A","#FFFF99", "#B15928")
# New color palette #
pal4 <- c("#1F78B4", "#FF7F00", "#33A02C", "#E31A1C")
# Order levels
carbon.cpds$Horizon <- factor(carbon.cpds$Horizon,levels = c("Organic","Mineral"))

# Plotting
p <- ggplot(carbon.cpds,aes(x=bs1_class,y=value,fill=Treatment))+ geom_boxplot()+
      stat_summary(fun.y=mean, geom="point", shape=20, size=3,position=position_dodge(width=0.75), color="red") +
      ylab("Percentage") + xlab("") + scale_fill_manual(values=pal4) + 
      theme(axis.text.x=element_text(size=12,face="bold",angle = 90,vjust = 0.7),
            axis.text.y=element_text(size=18,face="bold"),
            strip.text.x = element_text(size = 16),
            legend.text=element_text(size=18))
p + facet_grid(~Horizon)

```

```{r}
#stats on relative abundances #
results.temp<-ddply(carbon.cpds,.(bs1_class,Horizon),function(x) summarize(x,Pvalue=t.test(value~Treatment,data=x,na.rm=TRUE,paired=F)$p.value))
results.temp
```



# Ordination Plot:Treatment

```{r,fig.width=8,fig.height=4}
rm(list = ls())
load("ficusPeakObj.Robj")
peakObj <- group_designation(peakObj, main_effects=c("Treatment"))
pcoaMat <- getPrincipalCoordinates(ftmsObj = peakObj,dist_metric = "bray")
plotPrincipalCoordinates(pcoaMat,x=1,y=2,ftmsObj = peakObj )
```


# Ordination Plot:Treatment x Horizon

```{r,fig.width=8,fig.height=4}
peakObj <- group_designation(peakObj, main_effects=c("T_H"))
#getGroupDF(peakObj)
pcoaMat <- getPrincipalCoordinates(ftmsObj = peakObj,dist_metric = "bray")
plotPrincipalCoordinates(pcoaMat,x=1,y=2,ftmsObj = peakObj )
```

# Venny Diagram

```{r,fig.width=10}
rm(list = ls())
# load organic
load("organicUnique.Robj")
# get values
organic.unique.df <- organic_unique$e_data
# load mineral
load("mineralUnique.Robj")
# get values
mineral.unique.df <- mineral_unique$e_data
# Organic data
All.organic <- organic.unique.df[organic.unique.df$uniqueness_nsamps %in% "Observed in Both","Mass"]
# Mineral data
All.mineral <- mineral.unique.df[mineral.unique.df$uniqueness_nsamps %in% "Observed in Both","Mass"]
# Create List
venn.data <- list(
  Control.Mineral = c(mineral.unique.df[mineral.unique.df$uniqueness_nsamps %in% "Unique to Control","Mass"],All.mineral),
  Control.Organic = c(organic.unique.df[organic.unique.df$uniqueness_nsamps %in% "Unique to Control","Mass"],All.organic),
  Heated.Mineral = c(mineral.unique.df[mineral.unique.df$uniqueness_nsamps %in% "Unique to Heated","Mass"],All.mineral),
  Heated.Organic = c(organic.unique.df[organic.unique.df$uniqueness_nsamps %in% "Unique to Heated","Mass"],All.organic)
  )

# Venn plot
ggVennDiagram(venn.data,label_alpha = 0)
```




# Elements Plot


### Load Peak Object

```{r}
load("ficusPeakObj.Robj")
```


### Binary Matrix

```{r}
# Set rownames
edata <- peakObj$e_data
rownames(edata) <- as.character(edata$Mass)
edata$Mass <- NULL
# Binary matrix
edata.binary <- apply(edata,2,function(y)as.numeric(y>0))
# Na's as zero
edata.binary[is.na(edata.binary)] <- 0
# As dataframe edata.binary
edata.binary <- as.data.frame(edata.binary)
# Add Mass to binary matrix
edata.binary$Mass <- rownames(edata)
# Add metadata to binary matrix
tmp <- peakObj$e_meta[,c("Mass","ElComposition")] # save to tmp; direct merge isn't working
str(tmp)
# Merge 
edata.binary <- merge(edata.binary,tmp,by="Mass",all = T)
# Remove Mass column from df
edata.binary$Mass <- NULL
```


```{r}
# Carbon compounds #
carbon.cpds <- aggregate(.~ElComposition,data=edata.binary,FUN=sum)
# Transfor count to relative abundance #
carbon.cpds[,-1] <- decostand(carbon.cpds[,-1],method = "total",MARGIN = 2)*100
# Sum check
colSums(carbon.cpds[,-1])
```

```{r}
# melt df #
carbon.cpds <- melt(carbon.cpds,variable.name = "Sample_ID")
# Add sample info
carbon.cpds <- merge(carbon.cpds,peakObj$f_data,by="Sample_ID")
head(carbon.cpds)
```

## Elements Plot

```{r,fig.width=15}
# New color palette #
pal4 <- c("#1F78B4", "#FF7F00", "#33A02C", "#E31A1C")

# Order levels
carbon.cpds$Horizon <- factor(carbon.cpds$Horizon,levels = c("Organic","Mineral"))

# Plotting
p <- ggplot(carbon.cpds,aes(x=ElComposition,y=value,fill=Treatment))+ geom_boxplot()+
      stat_summary(fun.y=mean, geom="point", shape=20, size=5,position=position_dodge(width=0.75), color="red") +  
      ylab("Relative Frequency (%)") + xlab("") + scale_fill_manual(values=pal4) + 
      theme(axis.text.x=element_text(size=12,face="bold",angle = 90,vjust = 0.7),
            axis.text.y=element_text(size=18,face="bold"),
            strip.text.x = element_text(size = 16),
            legend.text=element_text(size=18))
p + facet_grid(~Horizon)

```

### Elements Stats

```{r}
#stats on relative abundances #
results.temp<-ddply(carbon.cpds,.(ElComposition,Horizon),function(x) summarize(x,Pvalue=t.test(value~Treatment,data=x,na.rm=TRUE,paired=F)$p.value))
results.temp
```



## Nitrogen Compounds

```{r}
# Filter by N-containing elements
tmp <- edata.binary[grep(pattern = "N",edata.binary$ElComposition),]
# Carbon compounds #
carbon.cpds <- aggregate(.~ElComposition,data=tmp,FUN=sum)
# Transfor count to relative abundance #
carbon.cpds[,-1] <- decostand(carbon.cpds[,-1],method = "total",MARGIN = 2)*100
# Sum check
colSums(carbon.cpds[,-1])
```

```{r}
# melt df #
carbon.cpds <- melt(carbon.cpds,variable.name = "Sample_ID")
# Add sample info
carbon.cpds <- merge(carbon.cpds,peakObj$f_data,by="Sample_ID")
head(carbon.cpds)
```

### Nitrogen Elements: Plot

```{r,fig.width=15}
# Order levels
carbon.cpds$Horizon <- factor(carbon.cpds$Horizon,levels = c("Organic","Mineral"))

# Plotting
p <- ggplot(carbon.cpds,aes(x=ElComposition,y=value,fill=Treatment))+ geom_boxplot()+
      stat_summary(fun.y=mean, geom="point", shape=20, size=5,position=position_dodge(width=0.75), color="red") +  
      ylab("Relative Frequency (%)") + xlab("") + scale_fill_manual(values=pal4) + 
      theme(axis.text.x=element_text(size=12,face="bold",angle = 90,vjust = 0.7),
            axis.text.y=element_text(size=18,face="bold"),
            strip.text.x = element_text(size = 16),
            legend.text=element_text(size=18))
p + facet_grid(~Horizon)

```

### Nitrogen Stats

```{r}
#stats on relative abundances #
results.temp<-ddply(carbon.cpds,.(ElComposition,Horizon),function(x) summarize(x,Pvalue=t.test(value~Treatment,data=x,na.rm=TRUE,paired=F)$p.value))
results.temp
```


## Phosphorus Compounds

```{r}
# Filter by N-containing elements
tmp <- edata.binary[grep(pattern = "P",edata.binary$ElComposition),]
# Carbon compounds #
carbon.cpds <- aggregate(.~ElComposition,data=tmp,FUN=sum)
# Transfor count to relative abundance #
carbon.cpds[,-1] <- decostand(carbon.cpds[,-1],method = "total",MARGIN = 2)*100
# Sum check
colSums(carbon.cpds[,-1])
```

```{r}
# melt df #
carbon.cpds <- melt(carbon.cpds,variable.name = "Sample_ID")
# Add sample info
carbon.cpds <- merge(carbon.cpds,peakObj$f_data,by="Sample_ID")
head(carbon.cpds)
```

### Phosphorus Elements: Plot

```{r,fig.width=15}
# Order levels
carbon.cpds$Horizon <- factor(carbon.cpds$Horizon,levels = c("Organic","Mineral"))

# Plotting
p <- ggplot(carbon.cpds,aes(x=ElComposition,y=value,fill=Treatment))+ geom_boxplot()+
      stat_summary(fun.y=mean, geom="point", shape=20, size=5,position=position_dodge(width=0.75), color="red") +
      ylab("Relative Frequency (%)") + xlab("") + scale_fill_manual(values=pal4) + 
      theme(axis.text.x=element_text(size=12,face="bold",angle = 90,vjust = 0.7),
            axis.text.y=element_text(size=18,face="bold"),
            strip.text.x = element_text(size = 16),
            legend.text=element_text(size=18))
p + facet_grid(~Horizon)

```

### Phosphorus Stats

```{r}
#stats on relative abundances #
results.temp<-ddply(carbon.cpds,.(ElComposition,Horizon),function(x) summarize(x,Pvalue=t.test(value~Treatment,data=x,na.rm=TRUE,paired=F)$p.value))
results.temp
```


## Sulfur Compounds

```{r}
# Filter by N-containing elements
tmp <- edata.binary[grep(pattern = "S",edata.binary$ElComposition),]
# Carbon compounds #
carbon.cpds <- aggregate(.~ElComposition,data=tmp,FUN=sum)
# Transfor count to relative abundance #
carbon.cpds[,-1] <- decostand(carbon.cpds[,-1],method = "total",MARGIN = 2)*100
# Sum check
colSums(carbon.cpds[,-1])
```

```{r}
# melt df #
carbon.cpds <- melt(carbon.cpds,variable.name = "Sample_ID")
# Add sample info
carbon.cpds <- merge(carbon.cpds,peakObj$f_data,by="Sample_ID")
head(carbon.cpds)
```

### Sulfur Elements: Plot

```{r,fig.width=15}
# Order levels
carbon.cpds$Horizon <- factor(carbon.cpds$Horizon,levels = c("Organic","Mineral"))

# Plotting
p <- ggplot(carbon.cpds,aes(x=ElComposition,y=value,fill=Treatment))+ geom_boxplot()+
      stat_summary(fun.y=mean, geom="point", shape=20, size=5,position=position_dodge(width=0.75), color="red") +
      ylab("Relative Frequency (%)") + xlab("") + scale_fill_manual(values=pal4) + 
      theme(axis.text.x=element_text(size=12,face="bold",angle = 90,vjust = 0.7),
            axis.text.y=element_text(size=18,face="bold"),
            strip.text.x = element_text(size = 16),
            legend.text=element_text(size=18))
p + facet_grid(~Horizon)

```

### Sulfur Stats

```{r}
#stats on relative abundances #
results.temp<-ddply(carbon.cpds,.(ElComposition,Horizon),function(x) summarize(x,Pvalue=t.test(value~Treatment,data=x,na.rm=TRUE,paired=F)$p.value))
results.temp
```


## Oxygen Compounds

```{r}
# Filter by N-containing elements
tmp <- edata.binary[grep(pattern = "O",edata.binary$ElComposition),]
# Carbon compounds #
carbon.cpds <- aggregate(.~ElComposition,data=tmp,FUN=sum)
# Transfor count to relative abundance #
carbon.cpds[,-1] <- decostand(carbon.cpds[,-1],method = "total",MARGIN = 2)*100
# Sum check
colSums(carbon.cpds[,-1])
```

```{r}
# melt df #
carbon.cpds <- melt(carbon.cpds,variable.name = "Sample_ID")
# Add sample info
carbon.cpds <- merge(carbon.cpds,peakObj$f_data,by="Sample_ID")
head(carbon.cpds)
```

### Oxygen Elements: Plot

```{r,fig.width=15}
# Order levels
carbon.cpds$Horizon <- factor(carbon.cpds$Horizon,levels = c("Organic","Mineral"))

# Plotting
p <- ggplot(carbon.cpds,aes(x=ElComposition,y=value,fill=Treatment))+ geom_boxplot()+
      stat_summary(fun.y=mean, geom="point", shape=20, size=5,position=position_dodge(width=0.75), color="red") +
      ylab("Relative Frequency (%)") + xlab("") + scale_fill_manual(values=pal4) + 
      theme(axis.text.x=element_text(size=12,face="bold",angle = 90,vjust = 0.7),
            axis.text.y=element_text(size=18,face="bold"),
            strip.text.x = element_text(size = 16),
            legend.text=element_text(size=18))
p + facet_grid(~Horizon)

```

### Oxygen Stats

```{r}
#stats on relative abundances #
results.temp<-ddply(carbon.cpds,.(ElComposition,Horizon),function(x) summarize(x,Pvalue=t.test(value~Treatment,data=x,na.rm=TRUE,paired=F)$p.value))
results.temp
```




